FP/TP/Dropped hours based on 6 minute review rules
Sum across t/n for all hours


HOUR:

3, .8; 2, .7; 4, .6; 1, .9

nSixDropped (TP & FP)
nSixReviewed (TP & FP)
nSixAccepted (TP & FP) (no review)
hourStatus - Accepted, Dropped, Reviewd
hourResult - FP, TP, FN

May need to add # of positive six minutes. Idea is maybe that there are hours where we have
dropped everything except for 1 or 2 that are false positives, and hopefully if its TP
we would have more? I Think? First pass doesnt seem to support this, but cant figure out
if I'm doing it the right way for a given set of nMin/revThresh

```{r}
nThreshSummary <- function(x, nMin=1:5, tComp=c(.7, .75, .8, .85, .9), evCol='eventId') {
    hasTruth <- 'evResult' %in% colnames(x)
    nSix <- nrow(x)
    nEvents <- length(unique(x[[evCol]]))
    eventSummary <- bind_rows(lapply(split(x, x[[evCol]]), function(e) {
        bind_rows(lapply(nMin, function(n) {
            dropIx <- e$nPreds < n
            thisSum <- e[!dropIx, ]
            result <- bind_rows(lapply(tComp, function(t) {
                revIx <- thisSum$meanPos <= t
                result <- list(nMin=n,
                               revThresh=t)
                if(hasTruth) {
                    result$tpReviewed <- sum(thisSum$evResult[revIx] == 'TP')
                    result$fpReviewed <- sum(thisSum$evResult[revIx] == 'FP')
                    result$tpAccepted <- sum(thisSum$evResult[!revIx] == 'TP')
                    result$fpAccepted <- sum(thisSum$evResult[!revIx] == 'FP')
                }
                result$reviewStatus <- 'Dropped'
                nAccepted <- sum(!revIx)
                nReviewed <- sum(revIx)
                result$sixAccepted <- nAccepted
                result$sixReviewed <- nReviewed
                if(nAccepted > 0) {
                    result$reviewStatus <- 'Accepted'
                } else if(nReviewed > 0) {
                    result$reviewStatus <- 'Review'
                }
                result
            }))
            if(hasTruth) {
                hourTrue <- any(e$finEvent)
                hourPred <- any(e$predEvent)
                tpDropped <- sum(e$evResult[dropIx] =='TP')
                fpDropped <- sum(e$evResult[dropIx] == 'FP')
                result$hourResult <- typeLab(hourTrue, hourPred)
                result$tpDropped <- tpDropped
                result$fpDropped <- fpDropped
            }
            result
        }))
    }), .id=evCol)
    threshSummary <- group_by(eventSummary, nMin, revThresh)
    if(hasTruth) {
        threshSummary <- summarise(threshSummary, 
                                   nDropped = sum(reviewStatus == 'Dropped'),
                                   nAccepted = sum(reviewStatus == 'Accepted'),
                                   nReviewed = sum(reviewStatus == 'Review'),
                                   tpSixReviewed=sum(tpReviewed * (reviewStatus == 'Review')),
                                   fpSixReviewed=sum(fpReviewed * (reviewStatus == 'Review')),
                                   sixReviewed = sum(sixReviewed * (reviewStatus == 'Review')),
                                   
                                   # tpSixAccepted=sum(tpAccepted),
                                   # fpSixAccepted=sum(fpAccepted),
                                   # tpSixDropped=sum(tpDropped),
                                   # fpSixDropped=sum(fpDropped),
                                   tpDropped = sum(reviewStatus == 'Dropped' & hourResult == 'TP'),
                                   tpAccepted = sum(reviewStatus == 'Accepted' & hourResult == 'TP'),
                                   tpReviewed = sum(reviewStatus == 'Review' & hourResult == 'TP'),
                                   fpDropped = sum(reviewStatus == 'Dropped' & hourResult == 'FP'),
                                   fpAccepted = sum(reviewStatus == 'Accepted' & hourResult == 'FP'),
                                   fpReviewed = sum(reviewStatus == 'Review' & hourResult == 'FP')
        )
    } else {
        threshSummary <- summarise(threshSummary,
                                   nDropped = sum(reviewStatus == 'Dropped'),
                                   nAccepted = sum(reviewStatus == 'Accepted'),
                                   nReviewed = sum(reviewStatus == 'Review'),
                                   sixReviewed = sum(sixReviewed * (reviewStatus == 'Review'))
        )
    }
    threshSummary$sixReviewed <- threshSummary$sixReviewed / nSix
    out <- list(eventSummary=eventSummary, threshSummary=threshSummary, nEvents=nEvents, nSix=nSix)
    if(hasTruth) {
        finSix <- sum(x$finEvent)
        finEvents <- group_by(x, .data[[evCol]]) %>% 
            summarise(hourEvent=any(finEvent))
        finEvents <- sum(finEvents$hourEvent)
        out$finEvents <- finEvents
        out$finSix <- finSix
    }
    out
}
hm <- nThreshSummary(filter(sixGroup, drift %in% len22))
```

```{r}
plotNThreshSummary <- function(x, nMin=1:5, tComp=c(.7, .75, .8, .85, .9), evCol='eventId') {
    if(is.data.frame(x)) {
        x <- nThreshSummary(x, nMin=nMin, tComp=tComp, evCol=evCol)
    }
    threshData <- x$threshSummary %>% 
        mutate(accText = paste0(tpAccepted, ' / ', fpAccepted, ' / ',
                                round((fpAccepted/x$finEvents), 2)),
               revText = paste0(tpSixReviewed, ' / ', fpSixReviewed, ' / ',
                                round(sixReviewed, 2)),
               dropText = paste0(tpDropped, ' / ', fpDropped))
    title <- paste0()
    acceptPlot <- ggplot() +
        geom_tile(data=threshData, aes(x=nMin, y=revThresh, fill=nAccepted)) +
        geom_text(data=threshData, aes(x=nMin, y=revThresh, label=accText)) +
        scale_fill_gradientn(colors=rev(scales::viridis_pal()(25))) +
        ggtitle(paste0('TP/FP Accepted (',x$finEvents, ' Fin & ', x$nEvents, ' Total Events)'))
    
    revPlot <- ggplot(threshData) +
        geom_tile(aes(x=nMin, y=revThresh, fill=nReviewed)) +
        geom_text(aes(x=nMin, y=revThresh, label=revText)) +
        scale_fill_gradientn(colors=rev(scales::viridis_pal()(25))) +
        ggtitle(paste0('TP/FP Reviewed (',x$finSix, ' Fin & ', x$nSix, ' Total Six)'))
    
    dropPlot <- ggplot(threshData, aes(x=nMin, y=revThresh)) +
        geom_tile(aes(fill=nDropped)) +
        geom_text(aes(label=dropText)) +
        scale_fill_gradientn(colors=rev(scales::viridis_pal()(25)))
    list(acceptPlot=acceptPlot, revPlot=revPlot, dropPlot=dropPlot)
}
plotNThreshSummary(hm)
```

Something about number of 6min files to review in an event series

```{r}
hm$eventSummary %>% 
    filter(nMin == 2 & revThresh == .85,
           reviewStatus == 'Reviewed') %>% 
    ggplot(aes(x=sixReviewed, fill=hourResult)) +
    geom_histogram()
```

sixGroup <- summarisePredEvent(newPdf, evCol=c('drift', 'eventId', 'sixMin'), thresh=.5)

```{r}
sixRevs <- markReviewStatus(allPdf, nMin=3, revThresh = .85, evCol=c('drift', 'hour', 'sixMin'))
table(sixRevs$reviewStatus)
```
```{r}
allPdf$hour <- floor_date(allPdf$UTC, unit = '1hour')
allPdf$sixMin <- floor_date(allPdf$UTC, unit='6mins')
allReview <- markReviewStatus(allPdf, nMin=3, revThresh = .85, evCol=c('drift', 'hour', 'sixMin'))
table(allReview$reviewStatus)
allReview <- allReview %>% 
    group_by(drift, hour) %>% 
    mutate(nReview = sum(reviewStatus == 'Review'),
           nAccept = sum(reviewStatus == 'Accepted'),
           hourStatus = ifelse(nReview > 0, 'Review', 'Dropped'),
           hourStatus = ifelse(nAccept > 0, 'Accepted', hourStatus)) %>% 
    ungroup()
```

```{r}


filter(allReview, drift == 'ADRIFT_028') %>% 
    split(day(.$hour)) %>%
    lapply(plotReviewStatus)
```

Notes for review process:
Create a folder of images for each drifts review events, and create a CSV in there
that is pre-filled with the review data. Then have an ingest function to read in that
CSV to mark events appropriately

IDEA for each drift make a grid of each row is an hour, column is 6 minute. Color
by outline is review status, fill is n preds or mean pos or something like that